---
title: "worms"
author: "Lynndsay Terpsma"
date: "2024-07-27"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(naniar)
library(skimr)
library(broom)
library(janitor)
library(googlesheets4)
library(stats)
library(jtools)
library(ggstance)
library(arm)
library(pscl)
library(lme4)
library(car)
library(fitdistrplus)
library(wesanderson)
```

##WITHOUT DIAMETER MEASUREMENTS
```{r load_data}
totalworms <- read_csv("C:/Users/lterp/Desktop/OneDrive - UBC/2024S/Project/Statistics/Data/MIZ_2024_worm_total.csv")
```
##VIEW DATA

```{r}
glimpse(totalworms)
```

##RECODE DATA - make fctr instead of chr, change names

```{r}
totalworms <- totalworms %>%
  mutate(testing_group = recode_factor(testing_group, 
                                       A1 = "A", A2 = "A", A3 = "A", 
                                       B1 = "B", B2 = "B", B3 = "B", 
                                       D1 = "D", D2 = "D", D3 = "D"))

totalworms
```
##Stripchart of data without tube diameter

```{r}
totalworms %>%
  ggplot(aes(x = testing_group, y = CR_CE_time)) +
  geom_jitter(colour = "darkgrey", size = 3, shape = 1, width = 0.1) +
  stat_summary(fun.data = mean_se, geom = "errorbar", 
    colour = "black", width = 0.1, 
    position = position_nudge(x = 0.15)) +
  stat_summary(fun = mean, geom = "point", 
    colour = "black", size = 3, 
    position = position_nudge(x = 0.15)) +
  xlab("Sediment treatment") + 
  ylab("Time to re-emerge (s)") + 
  theme_classic()
```
##normality check (data without tube diameter)

```{r}
totalworms %>% 
 ggplot(aes(sample = CR_CE_time)) +
  stat_qq(shape = 1, size = 2) +
  stat_qq_line() +
  ylab("re-emergence") +
  xlab("Normal quantile") +
  theme_bw()
```
##normality check (without diameter)

```{r}
hist(totalworms$CR_CE_time)
```
##LN TRANSFORMED DATA

```{r}
logtotalworms <- totalworms %>%
  mutate(logCR_CE_time_1 = log(CR_CE_time + 1))
```

```{r}
hist(logtotalworms$logCR_CE_time_1)
```
##LOG 10 TRANSFORMED DATA

```{r}
log10totalworms <- totalworms %>%
  mutate(log10CR_CE_time_1 = log10(CR_CE_time + 1))
log10totalworms
```

```{r}
hist(log10totalworms$log10CR_CE_time_1)
```

```{r}
log10totalworms_clean <- log10totalworms %>%
  filter(!is.na(log10CR_CE_time_1))
fit_norm <- fitdist(log10totalworms_clean$log10CR_CE_time_1, "norm")
fit_exp <- fitdist(log10totalworms_clean$log10CR_CE_time_1, "exp")
```

```{r}
gof <- gofstat(list(fit_norm, fit_exp))
print(gof)
```

```{r}
plot.legend <- c("normal", "exp")
denscomp(list(fit_norm, fit_exp), legendtext = plot.legend)
cdfcomp(list(fit_norm, fit_exp))
qqcomp(list(fit_norm, fit_exp))
ppcomp(list(fit_norm, fit_exp))
```

##ORIGINAL DATA (NOT TRANSFORMED) - without diameter MODEL SELECTION


###model 1
```{r model}
model <- glmer(CR_CE_time ~ testing_group + (1|worm_code/trial), data = totalworms, family = poisson)
summary(model)
confint(model)
```

```{r}
plot(residuals(model))
```

```{r}
Anova(model)
```

###model2
```{r}
model2 <- glmer(CR_CE_time ~ testing_group + (1|trial), data = totalworms, family = poisson)
summary(model2)
```
###model3
```{r}
model3 <- glmer(CR_CE_time ~ testing_group + (1|worm_code), data = totalworms, family = poisson)
summary(model)
```
###model comparison
```{r}
MuMIn::model.sel(model, model2, model3)
```

##Load data with diameter
```{r}
wormworm <- read_csv("C:/Users/lterp/Desktop/OneDrive - UBC/2024S/Project/Statistics/Data/MIZ_2024_Worm_Data_total2.csv")
```
##recode data with diameter
```{r}
wormworm <- wormworm %>%
  mutate(testing_group = recode_factor(testing_group, 
                                       A1 = "Control", A2 = "Control", A3 = "Control", 
                                       B1 = "Moderate", B2 = "Moderate", B3 = "Moderate", 
                                       D1 = "High", D2 = "High", D3 = "High"))

wormworm
```
##MODEL COMPARISON WITH DIAMETER

###model1 - this is the one we reported in the project
```{r tube_diameter}
modell1 <- glmer(CR_CE_time ~ testing_group + tube_diameter_cm + (1|worm_code/trial), data = wormworm, family = poisson)
summary(modell1)
confint(modell1)
```
```{r}
Anova(modell1)
```

###estimates of modell1
```{r estimate_df}
estimates <- data.frame(
  testing_group = c("Control", "Moderate", "High"),
  Estimate = c(4.7231, 4.7231 + (-0.6455), 4.7231 + (-0.6109)),
  CI_lower = c(4.1838535, 4.7231 + (-1.0326556), 4.7231 + (-1.0108977)),
  CI_upper = c(5.1910929, 4.7231 + (-0.2535972), 4.7231 + (-0.2071626))
)
```

we ran a mixed effects model with random effects for (...), because not normal distribution we specified our distribution (poisson). control from model: 4.7231 [95% CI: 4.1838535 - 5.2517116], B: estimate value [95 CI: 4.7231 + (-1.0326556) - 4.7231 + (-1.0108977)]

```{r}
print(estimates)
```

###anova of modell1
```{r}
anova_results <- Anova(modell1)
anova_results
```
Anova results - report both. talk about units (seconds)

```{r}
anova_df <- as.data.frame(anova_results)
anova_df$Term <- rownames(anova_df)
anova_df <- anova_df[, c("Term", "Chisq", "Df", "Pr(>Chisq)")]
rownames(anova_df) <- NULL
```

```{r, include = FALSE, eval = FALSE}
anova_ft <- flextable(anova_df)
anova_ft <- set_caption(anova_ft, "Type II Wald Chi-square ANOVA Results")
anova_ft <- theme_vanilla(anova_ft)
anova_ft <- autofit(anova_ft)
print(anova_ft)
```

###violinplot of modell1

```{r}
palette <- wes_palette("Darjeeling1")
```


```{r violin_plot}
  ggplot() +
  geom_violin(data = wormworm, aes(x = testing_group, y = log(CR_CE_time), fill = testing_group)) +
  scale_fill_manual(values = palette) +
  geom_point(data = estimates, aes(x = testing_group, y = Estimate), colour = "black", size = 2) +
  geom_errorbar(data = estimates, aes(x = testing_group, ymin = CI_lower, ymax = CI_upper), width = 0.1, colour = "black") +
  xlab("Sediment Treatment Level") +
  ylab("Log (ln) Transformed Time to Re-Emerge (s)") +
  labs(fill = "Sediment Level") +
  theme_bw()
```

###model comparison between with and without diameter

```{r}
MuMIn::model.sel(model, modell1)
```
##does trial number effect re-emergence?

```{r trial}
trialmodel <- glmer(CR_CE_time ~ testing_group + tube_diameter_cm + trial + (1|worm_code), data = wormworm, family = poisson)
summary(trialmodel)
confint(trialmodel)
```
```{r tube_diameter_histogram}
ggplot(wormworm, aes(x = tube_diameter_cm)) +
  geom_histogram(colour = "black", fill = "lightgrey") +
  labs(x = "Tube Diameter (cm)", y = "Frequency") +
  theme_bw()
```




#Partial Retraction

```{r PR_stats}
PR.stats <- wormworm %>%
  group_by(testing_group) %>%
  summarise(
    Count_PR = n() - naniar::n_miss(PR_Time),
    .groups = 'drop'
  ) %>%
    mutate(
      Total_count_PR = sum(Count_PR),
      Relative_freq_PR = Count_PR / Total_count_PR
  ) 

kable(PR.stats, digits = 4)
```
```{r PR_table}
PR.table <- wormworm %>%
  group_by(testing_group) %>%
  summarise(
    Count_PR = n() - naniar::n_miss(PR_Time),
    .groups = 'drop'
  ) %>%
  mutate(
    Total_count_PR = sum(Count_PR),
    Relative_freq_PR = Count_PR / Total_count_PR
  )
print(PR.table)
```

```{r PR_sort}
PR.stats <- wormworm %>%
  group_by(testing_group) %>%
  summarise(
    Count_PR = n() - naniar::n_miss(PR_Time),
    .groups = 'drop'
  ) 
```


##Partial retraction contingency table and chi-square test
```{r}
contingency_tab_wide <- PR.stats %>%
  pivot_wider(names_from = testing_group, values_from = Count_PR, values_fill = list(Count_PR = 0))
print(contingency_tab_wide)
```
```{r}
contingency_matrix <- as.matrix(contingency_tab_wide)
print(contingency_matrix)
```
```{r}
chi_square <- chisq.test(contingency_matrix)
print(chi_square)
```




#Partial Re-emergence

```{r}
PE.stats <- wormworm %>%
  group_by(testing_group) %>%
  summarise(
    Count_PE = n() - naniar::n_miss(PE_Time),
    .groups = 'drop'
    ) %>%
    mutate(
      Total_count_PE = sum(Count_PE),
      Relative_freq_PE = Count_PE / Total_count_PE
  ) 

kable(PE.stats, digits = 4)
```
##Exposed

```{r}
EX.stats <- wormworm %>%
  group_by(testing_group) %>%
  summarise(
    Count_EX = n() - naniar::n_miss(EX),
    .groups = 'drop'
    ) %>%
    mutate(
      Total_count_EX = sum(Count_EX),
      Relative_freq_EX = Count_EX / Total_count_EX
  ) 

kable(EX.stats, digits = 4)
```
##Hidden

```{r}
HI.stats <- wormworm %>%
  group_by(testing_group) %>%
  summarise(
    Count_HI = n() - naniar::n_miss(HI),
    .groups = 'drop'
    ) %>%
    mutate(
      Total_count_HI = sum(Count_HI),
      Relative_freq_HI = Count_HI / Total_count_HI
  ) 

kable(HI.stats, digits = 4)
```











